---
title: "p8105_hw1_zl3386"
author: "Ziqiu Liu"
date: "2023-09-14"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(moderndive)
data("early_january_weather")
#help(early_january_weather)
```

# Problem 1

## The `early_january_weather` dataset

The `early_january_weather` dataset contains hourly meterological data for LGA, JFK and EWR for the month of January 2013. It is a subset of the `weather` data frame from `nycflights13`.

* It has 15 variables, including `origin`(weather station), `year`, `month`, `day`, `hour`, `time_hour`(date and hour of the recording as a _POSIXct_ date) and indicators of weather.
* It has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.
* The mean temperature is `r mean(early_january_weather$temp)` °F.

## Scatterplot

Using the data from`early_january_weather`, we can make a scatterplot of `temp` vs `time_hour`, using `humid` to determine the color of each point.

```{r df_create, echo = FALSE}
plot_df = tibble(
  x = early_january_weather$time_hour,
  y1 = early_january_weather$temp,
  humidity = early_january_weather$humid
)
```

```{r yx_scatter, echo = FALSE}
ggplot(plot_df, aes(x = x, y = y1, color = humidity)) + geom_point()

```

From the scatterplot, we can see that the temperature in the month of January 2013 has fluctuations but goes up in general over time, and the relative humidity is higher in late January.

We will save this scatterplot as a png file (`scatterplot.png`) in the project directory.

```{r save plot, echo = FALSE, message = FALSE}
ggsave("scatterplot.png")
```

# Problem 2

## Create a dataframe
For Problem 2, we will create a data frame `example_df` comprised of:

* `num_samp`: a random sample of size 10 from a standard Normal distribution.
* `log_vec`: a logical vector indicating whether elements of the sample are greater than 0.
* `char_vec`: a character vector of length 10.
* `fac_vec`: a factor vector of length 10, with 3 different factor “levels”.

```{r create df}
example_df =
  tibble(
    num_samp = rnorm(10),
    log_vec = num_samp >=0,
    char_vec = c("M", "S", "S", "M", "L", "M", "M", "M", "S", "M"),
    fac_vec = factor(char_vec, levels= c("s", "M", "L"))
  )
```

## Take the mean
We can then try to take the mean of each variable in `example_df`.

```{r taking mean, warning = FALSE}
x1 = mean(pull(example_df, var = num_samp))
x2 = mean(pull(example_df, var = log_vec))
x3 = mean(pull(example_df, var = char_vec))
x4 = mean(pull(example_df, var = fac_vec))
```

* For `num_samp`, the mean is `r x1`.
* For `log_vec`, the mean is `r x2`.
* For `char_vec` and `fac_vec`, their mean cannot be taken and the return value is `r x3`.

## Explicitly convert variable types

Finally, we will try to convert variables from other types (logical, character and factor) to numeric.

```{r type conversion, message = FALSE, warning = FALSE}
y1 = as.numeric(pull(example_df, var = log_vec))
y2 = as.numeric(pull(example_df, var = char_vec))
y3 = as.numeric(pull(example_df, var = fac_vec))
```

* For `log_vec`, the logical value TRUE is converted to 1, and FALSE is converted to 0.
* For `char_vec`, all values are converted to NA.
* For `fac_vec`, the value of the first level is converted to NA, while the values of the other levels are converted according to their order.

This helps explain what happens when we try to take the mean -- when asked to calculate the mean of a non-numeric variable, R will first try to convert it to a numeric variable. But because of the NA that exists after the conversion, R cannot take the mean of character or factor variables, and can only return NA as a result.